\Large \bf Uncovering Distributed Storage System Bugs during\\Testing (not in Production!)

### 2015-09-21 methodology.tex

\textbf{Internals of the real Extent Manager}.
Inside the real ExtMgr (see Figure~\ref{fig:extentmanager}), there are two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} maintains the mapping records from extents to their hosting ENs. It is updated upon the periodic sync reports from the ENs. Recall that the sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update extent manager's possible out-of-date view of the EN with the ground truth. The EN map records the latest heartbeat time from every EN.

The ExtMgr runs internally a periodic {\em EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period, as well as cleaning up the corresponding records in the \texttt{ExtentCenter}. In addition, the ExtMgr runs a periodic {\em extent repair loop} that examines all the \texttt{ExtentCenter} records, identifies extents with missing replicas, schedules extent repair tasks and sends them to the ENs.

@@

Messages coming from the ENs do \emph{not} go through the mocked network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the internal ExtMgr with \texttt{extMgr.ProcessSingleMessage}. The benefit of this approach is that ExtMgr can be tested without modifying its code; ExtMgr is simply unaware of the testing harness and behaves as if it running in a real distributed environment and communicating with real ENs.

@@

  // extent repair logic
  ...
  [OnEvent(ExtentCopyResponse, ProcessExtentCopyResponse)]
  void ProcessExtentCopyResponse() {
    // extent copy response from source replica
    if (IsCopySucceeded(this.Payload)) {
      var rec = GetExtentRecord(this.Payload);
      extCtr.AddOrUpdate(rec); // update Extent Center
    }
  }

@@

The \texttt{RepairMonitor} machine is a \psharp liveness monitor (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state. Whenever an extent node fails, \texttt{RepairMonitor} is notified. As soon as the number of extent replicas falls below a specified target (three replicas in the current test harness), \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired. Whenever a replica is repaired, the \texttt{RepairMonitor} machine is also notified. It transitions into the cold \emph{repaired} state, when the replica number reaches again the target, as illustrated in Figure~\ref{fig:monitor}.

In the extent repair testing scenarios, \texttt{RepairMonitor} checks that it should {\em always eventually} end up in the cold repaired state. Otherwise, the \texttt{RepairMonitor} machine is stuck in the hot repairing state for {\em infinitely} long. This indicates that the corresponding execution sequence results in an extent replica never being repaired, which is a liveness bug.

@@

The liveness bug occurs in the second testing scenario, where the \texttt{TestingDriver} fails one of the \texttt{ExtentNode} and launches a new one. The \texttt{RepairMonitor} transitions to the hot repairing state and is stuck in the state for infinitely long.

The following is one particular execution sequence resulting in this liveness bug: (i) EN$_0$ fails and is detected by the EN expiration loop; (ii) EN$_0$ is removed from the EN map; (iii) the extent center is updated and the replica count drops from 3 (which is the target) to 2; (iv) ExtMgr receives a sync report from EN$_0$; (v) the extent center is updated and the replica count increases again from 2 to 3. This is problematic since the replica count is equal to the target, which means that the extent repair loop will never schedule any repair task. At the same time, there are only two \emph{true} replicas in the system, which is one less than the target. This execution sequence leads to one missing replica; repeating this process two more times would result in all replicas missing, but ExtMgr still thinking that all replicas are healthy. In production, such a bug can cause a very serious incident of customer data loss.

>>>>

\textbf{Internals of the real Extent Manager}.
Inside the real ExtMgr (see Figure~\ref{fig:extentmanager}), there are two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} data structure maintains the mapping records from extents to their hosting ENs. It is updated upon the periodic sync reports from the ENs. Recall that the sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update extent manager's possible out-of-date view of the EN with the ground truth. \texttt{ExtentNodeMap} records the latest heartbeat time from every EN.

The ExtMgr runs internally a periodic \emph{EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period of time, as well as cleaning up the corresponding extent records in \texttt{ExtentCenter}. In addition, the ExtMgr runs a periodic \emph{extent repair loop} that examines all the \texttt{ExtentCenter} records, identifies extents with missing replicas, schedules extent repair tasks and sends them to the ENs.

@@

Messages coming from the ENs do \emph{not} go through the mocked network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the internal ExtMgr with \texttt{extMgr.ProcessMessage}. The benefit of this approach is that ExtMgr can be tested without modifying its code; ExtMgr is simply unaware of the testing harness and behaves as if it running in a real distributed environment and communicating with real ENs.

@@

  // extent repair logic
  ...
  [OnEvent(ExtentCopyResponse, ProcessExtentCopyResponse)]
  void ProcessExtentCopyResponse() {
    // extent copy response from source replica
    if (IsCopySucceeded(this.Payload)) {
      var rec = GetExtentRecord(this.Payload);
      extCtr.AddOrUpdate(rec); // update ExtentCenter
    }
  }

@@

The \texttt{RepairMonitor} machine is a \psharp liveness monitor (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state. Whenever an EN fails, \texttt{RepairMonitor} is notified. As soon as the number of extent replicas falls below a specified target (three replicas in the current \psharp test harness), \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired. Whenever a replica is repaired, \texttt{RepairMonitor} is notified. It transitions into the cold \emph{repaired} state, when the replica number reaches again the target, as illustrated in Figure~\ref{fig:monitor}.

In the extent repair testing scenarios, \texttt{RepairMonitor} checks that it should \emph{always eventually} end up in the cold state. Otherwise, \texttt{RepairMonitor} is stuck in the hot state for \emph{infinitely} long. This indicates that the corresponding execution sequence results in an extent replica never being repaired, which is a liveness bug.

@@

The liveness bug occurs in the second testing scenario, where the \texttt{TestingDriver} machine fails one of the \texttt{ExtentNode} machines and launches a new one. \texttt{RepairMonitor} transitions to the hot repairing state and is stuck in the state for infinitely long.

The following is one particular execution sequence resulting in this liveness bug: (i) EN$_0$ fails and is detected by the EN expiration loop; (ii) EN$_0$ is removed from \texttt{ExtentNodeMap}; (iii) \texttt{ExtentCenter} is updated and the replica count drops from 3 (which is the target) to 2; (iv) ExtMgr receives a sync report from EN$_0$; (v) the extent center is updated and the replica count increases again from 2 to 3. This is problematic since the replica count is equal to the target, which means that the extent repair loop will never schedule any repair task. At the same time, there are only two \emph{true} replicas in the system, which is one less than the target. This execution sequence leads to one missing replica; repeating this process two more times would result in all replicas missing, but ExtMgr still thinking that all replicas are healthy. In production, such a bug can cause a very serious incident of customer data loss.
